---
title: "Correlations"
author: "Javier Porras"
date: "`r Sys.Date()`"
output: html_document
bibliography: references.bib
csl: citation_style/ecology.csl
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, error = FALSE, message = FALSE)
```

### Context

The non-destructive methodology we are currently following (based on @Jabbot) had not been yet decided in the first 3 samplings (sampling 0, 1 and 2). While now we are taking data of height ($H$), circumference at base level ($Cb$) and circumference at medium height level ($Ch$), in the 3 first samplings we only gathered information about $H$. To apply the models proposed by @Jabbot and @perronne2020 we need the 3 morphological measurements. Hence, we have to find a way to fill these gaps of data. 

### Approach

So far, excluding these 3 samplings, we have around 3 thousand data points, which include a measurement of each of the 3 morphological data for a certain individual from a species. Hence, we can use these data to check if there is any significant correlation between $H$ and $Cb$ and $Cm$. To do this, we can create linear models between these parameters per species and use $R-squared$ and $p-value$ as statistical indicators to check the significance of correlations. Therefore, we would have: 
- lm($Cb$ ~ $H$, flora). Where we can see the correlation between $H$ and $Cb$. 
- lm($Cm$ ~ $H$, flora). Where we can see the correlation between $H$ and $Cm$. 


```{r packages}
library(stringr)
library(dplyr)
library(reshape2)
library(readr)
library(ggpubr)
library(tidyverse)
library(ggplot2)
library(gridExtra)
```


```{r opening data}

flora_raw<- read.csv("data/flora_db.csv")
flora_raw <- flora_raw %>%
  mutate(across(where(is.character), as.factor))
flora_raw$Dm <- flora_raw$Dm + 0.01
flora_raw$Db <- flora_raw$Db + 0.01
pi <- 3.1416
flora_raw$cm <- round(ifelse(!is.na(flora_raw$Dm), flora_raw$Dm * pi, flora_raw$Cm), 2)
flora_raw$cb <- round(ifelse(!is.na(flora_raw$Db), flora_raw$Db * pi, flora_raw$Cb), 2)
flora <- flora_raw %>% select("species", "sampling", "height", "cb", "cm")

#List of species from samplings 0, 1 and 2
selected_samplings0 <- c("s_0", "s_1", "s_2")
flora_s012 <- flora[flora$sampling %in% selected_samplings0, ]

```

After filtering data, we have `r length(unique(flora_s012$species))` species to be subjected to the linear regression. We will choose only the data points stored in the database for these species. 

```{r Filtering data}

#adjustment for levels
flora_s012 <- droplevels(flora_s012)
#Creating database with species that are present on s_ 0,1 and 2. 
flora_s012_listofspecies <- unique(flora_s012$species)
species_s012 <- expand.grid(flora_s012_listofspecies)
colnames(species_s012) <- c("species")
#Extracting rows from "flora" that matches with the species found within flora_s012
flora_species_s012 <- merge(flora, species_s012)
#Choosing the rest of samplings, since in s_0, s_1 and s_2 are only NA values. 
selected_samplings <- c("s_3", "s_4", "s_5", "s_6", "s_7", "s_8", "s_9", "s_10", "s_11")
flora_morph_data <- flora_species_s012[flora_species_s012$sampling %in% selected_samplings, ]

#We have a lower n of species within the rest of samplings thant in the gap samplings. How many and which ones?
flora_species_s012$species <- droplevels(flora_species_s012$species)
flora_morph_data$species <- droplevels(flora_morph_data$species)
sp1 <- unique(flora_species_s012$species)
sp2 <- unique(flora_morph_data$species)
sp3 <- setdiff(sp1, sp2)

```

However, there are species that were only present during the first three samplings. The final number of species of which we have available data are **`r length(unique(flora_morph_data$species))`**. The lacking `r length(unique(flora_s012$species)) - length(unique(flora_morph_data$species))` species are the following ones: 
`r unique(sp3)`

### Correlations

As a result of applying the linear model regressions between $H$ and $Cb$ and $Cm$, we have found that barely 3 species have a significant linear correlation. Besides, only 1 out of these 3 species have a substantial amount of measurements included (Table 1.)

```{r Correlations}
# Count the number of rows for each species and delete those that have less than 2 rows (2 measurements) 
species_lower2<- flora_morph_data %>%
  group_by(species) %>%
  summarise(row_count = n()) %>%
  filter(row_count < 2)

flora_morph_data <- anti_join(flora_morph_data, species_lower2, by = "species")

#Cleaning NA's
flora_morph_data<- flora_morph_data %>% filter(!is.na(height) & !is.na(cb) & !is.na(cm))
library(purrr)

# CB vs HEIGHT #####
# Define a function to extract information for a given species
cb_df_function <- function(species) {
  subset_data <- filter(flora_morph_data, species == !!species)
  
  # Check if there are non-missing values for both height and cb
  if (sum(!is.na(subset_data$height) & !is.na(subset_data$cb)) > 0) {
    lm_model <- lm(cb ~ height, data = subset_data)
    
    return(data.frame(
      species = species,
      cb_R_squared = summary(lm_model)$r.squared,
      cb_P_value = summary(lm_model)$coefficients[2, 4],
      num_measurements = nrow(subset_data)
    ))
  } else {
    # If there are no non-missing values, return NA or any other indicator
    return(data.frame(
      species = species,
      cb_R_squared = NA,
      cb_P_value = NA, 
      num_measurements = 0
    ))
  }
}

# Apply the function to each species level and bind the results
cb_df_correlation <- map_dfr(sp2, cb_df_function)

# CM #####

# Define a function to extract information for a given species
cm_df_function <- function(species) {
  subset_data <- filter(flora_morph_data, species == !!species)
  
  # Check if there are non-missing values for both height and cm
  if (sum(!is.na(subset_data$height) & !is.na(subset_data$cm)) > 0) {
    lm_model <- lm(cm ~ height, data = subset_data)
    
    return(data.frame(
      species = species,
      cm_R_squared = summary(lm_model)$r.squared,
      cm_P_value = summary(lm_model)$coefficients[2, 4],
      num_measurements = nrow(subset_data)
    ))
  } else {
    # If there are no non-missing values, return NA or any other indicator
    return(data.frame(
      species = species,
      cm_R_squared = NA,
      cm_P_value = NA, 
      num_measurements = 0
    ))
  }
}

# Apply the function to each species level and bind the results
cm_df_correlation <- map_dfr(sp2, cm_df_function)


df_c_correlations <- merge(cb_df_correlation, cm_df_correlation)

library(knitr)
df_c_correlations %>%
  arrange(desc(cb_R_squared)) %>%
  slice_head(n = 7) %>%
  kable()

```


*Table 1. Top 7 Correlations based on R-squared values for $Cb$*


Since we have found no proper correlations for $Cb$ or $Cm$, we decided to try with the expression $X$. $X$ stands for the part of the equation provided by @perronne2020 that aggregates the three morphological values: $H$, $Cb$ and $Cm$. The circumferences are transformed in *area* values. 

So we have the general equation: 

$$
B = d \left [ \frac{H}{2} \left ( A_{basal} + A_{half} \right ) \right ]^{z},
$$
Where X should be: 

$$
x = \left [ \frac{H}{2} \left ( A_{basal} + A_{half} \right ) \right ],
$$ 

Therefore: 

$$
B = dx^{z}
$$
*d and z are respectively 1.96 and 2/3 (see @perronne2020)*

As a result, we have found an improvement in the number of species where there is an acceptable linear correlation (R-squared > 0.5). There is a total of 7 of these species, as we can see in Table 2. 

```{r Correlations for X}
#Añado el valor de X (modelo de la ecuacion de biomasa)

flora_morph_data$Ah <- ((flora_morph_data$cm)^2)/4*pi
flora_morph_data$Ab <- ((flora_morph_data$cb)^2)/4*pi

d <- 1.96
z <- 2/3
flora_morph_data$x <- (flora_morph_data$height/2)*(flora_morph_data$Ab + flora_morph_data$Ah)

# Define a function to extract information for a given species

x_df_function <- function(species) {
  subset_data <- filter(flora_morph_data, species == !!species)
  
  # Check if there are non-missing values for both height and cm
  if (sum(!is.na(subset_data$height) & !is.na(subset_data$x)) > 0) {
    lm_model <- lm(x ~ height, data = subset_data)
    
    return(data.frame(
      species = species,
      x_R_squared = summary(lm_model)$r.squared,
      x_P_value = summary(lm_model)$coefficients[2, 4],
      num_measurements = nrow(subset_data)
    ))
  } else {
    # If there are no non-missing values, return NA or any other indicator
    return(data.frame(
      species = species,
      x_R_squared = NA,
      x_P_value = NA, 
      num_measurements = 0
    ))
  }
}

# Apply the function to each species level and bind the results
x_df_correlation <- map_dfr(sp2, x_df_function)

x_df_correlation %>%
  arrange(desc(x_R_squared)) %>%
  slice_head(n = 10) %>%
  kable()

```

 *Table 2. Top 10 Correlations based on R-squared values for $X$*


### Further questions and tasks

- Can we do any "gap-filling" with these data?
- If we cannot, how should we proceed with the first 3 samplings?


### References

::: {#refs}
:::

<br>
