---
title: "Sensitivity analysis of model parameters"
date: "`r Sys.Date()`"
output: html_document
bibliography: references.bib
csl: citation_style/ecology.csl
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, error = FALSE, message = FALSE)
```

### Model

The allometric model described by is the following:

$$
B = d \left [ \frac{H}{2} \left ( A_{basal} + A_{half} \right ) \right ]^{z},
$$ where $B$ is the biomass calculated by the allometric equation, $d$ is the bulk density of the plant species, $H$ is the measured height of the plant, $A_{basal}$ and $A_{half}$ are the basal area and the area at half the height of the plant, respectively; and $z$ is the power coefficient of a power law.

### Problem

In @perronne2020, the authors assign a value of 1.96 to $d$ and a value of $2/3$ to $z$; however, we do not know how those parameters interact with each other.

### Approach

To solve this, we perform a sensitivity analysis in which we modify the values of $d$ and $z$ within a certain range, described below. Importantly, to simplify the allometric equation and do a clear sensitivity analyses that works independently of the field data, we opt to include the parameter $x$ as:

$$
x = \left [ \frac{H}{2} \left ( A_{basal} + A_{half} \right ) \right ],
$$ meaning that all allometric measurements are summarized now by $x$. Thus, the equation used for our sensitivity analysis is the following:

$$
B = dx^{z}
$$

```{r load_x}
x_data <- read.table("data/x_values.txt", header = TRUE, sep = "\t")
```

At this point, we assign the following values to the different parameters in the model:

```{r values}
d <- seq(0.98, 3.92, 0.49)
x <- seq(round(min(x_data$x), 0), round(max(x_data$x), 0), 25)
z <- seq(0.1, 1, 0.01)
```

-   for $d$, values of `r d`;
-   for $x$, values of `r x`;
-   for $z$, values of `r z`.

```{r df}
df <- expand.grid(d, x, z)
colnames(df) <- c("d", "x", "z")
df$biomass <- df$d * (df$x ^ df$z)
```

And we create a data frame including all the combinations between the values of each parameter (`r nrow(df)` different cases) and their predicted value of biomass, $B$.

### Dataviz

Now, some visualizations to understand the relationship between the different parameters.

```{r libraries}
library(tidyverse)
library(ggpubr)
```

#### $x$ and $z$

The relationship between $x$ and $z$ for values of $d$ = `r min(df$d)`, $d$ = 1.96, and $d$ = `r max(df$d)`, looks like this:

```{r tilexz}
dvalue <- 0.98
tilexz <- 
  ggplot(df[which(df$d == dvalue),], aes(x = x, y = z)) +
  geom_tile(aes(fill = biomass)) +
  labs(fill = "Biomass", title = paste("d", "=", dvalue, sep = " ")) +
  scale_x_continuous(expand = c(0, 0)) +
  scale_y_continuous(expand = c(0, 0)) +
  theme_bw()
```

```{r tilexz_mind}
dvalue <- 1.96
tilexz_mind <- 
  ggplot(df[which(df$d == dvalue),], aes(x = x, y = z)) +
  geom_tile(aes(fill = biomass)) +
  labs(fill = "Biomass", title = paste("d", "=", dvalue, sep = " ")) +
  scale_x_continuous(expand = c(0, 0)) +
  scale_y_continuous(expand = c(0, 0)) +
  theme_bw()
```

```{r tilexz_maxd}
dvalue <- 3.92
tilexz_maxd <- 
  ggplot(df[which(df$d == dvalue),], aes(x = x, y = z)) +
  geom_tile(aes(fill = biomass)) +
  labs(fill = "Biomass", title = paste("d", "=", dvalue, sep = " ")) +
  scale_x_continuous(expand = c(0, 0)) +
  scale_y_continuous(expand = c(0, 0)) +
  theme_bw()
```

```{r arrange_d}
ggarrange(tilexz, tilexz_mind, tilexz_maxd)
```

(...)

```{r d_biomass}
dvalue <- 1.96
ggplot(df[which(df$d == dvalue),], aes(x = z, y = biomass, color = as.factor(x))) +
  labs(color = "x", title = paste("d", "=", dvalue, sep = " ")) +
  geom_vline(xintercept = 2/3, linetype = "dashed") +
  geom_point() +
  theme_classic()
```

(...)

### References

::: {#refs}
:::

<br>
