---
title: "Plots"
author: "Javier Porras"
date: "`r Sys.Date()`"
output:
 
  html_document: default
  
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, error = FALSE, message = FALSE)
```

```{r, inclue = FALSE}
# to do's: 

# Include equations of logRR, CV, etc..

```


```{r package, include=FALSE}

library(tidyverse) # manage data
library(DT) # visualise tables in html
library(viridis) # use color-blind friendly palettes
library(ggrepel) # avoid overlapping texts in the figures
library(codyn) # calculate community dynamics metrics
library(vegan) # calculate community ecology metrics
library(ggpubr)
library(stringr)
library(ggplot2)
library(dplyr)
library(reshape2)
library(readr)
library(ggpubr)
library(tidyverse)
library(gridExtra)

```


``` {r Scripts}
source("code/first_script.R")
source("code/flora_bio_div_ab.R")
source("code/species_composition_analysis.R")

theme_set(theme_bw()+ theme(legend.position = "NULL"))

```

### Plots


Dynamics of abundance and richness

```{r}
ggDynamics
```





Coefficient of variation (CV) dynamics of abundance and richness

```{r}
ggCVgrid
```

Dynamics of evenness index from Rank Abundance Distribution (RADs) models (zipf and lognormal)

```{r}
ggDynamics_evenness
```

CV of evenness index

```{r}
ggCVgrid_evenness
```


Differences at sampling 0

```{r}
ggS0
```

Response Ratio (RR) of abundance and richness with control as reference

```{r}
ggRRcontrol
```

RR(CV abundance and richness) with control as reference

```{r}
ggRRcv

```

RR(evenness) with control as reference

```{r}
ggRRcontrol_evenness
```

RR(CV eveness) with control as reference

```{r}
ggRRcv_eveness
```

RR(abundance and richness). P and W as references of WP

```{r}
ggRRwp
```

RR(CV abundance and richness): P and W as references of WP

```{r}
ggRRcv_wp
```

RR(eveness): P and W as references of WP

```{r}
ggRRwp_evenness
```

RR(CV eveness): P and W as references of WP

```{r}
ggRRcv_wp_evenness

```

Species turnover

```{r}
ggturnover
```

PCoA using Hellinger distance

```{r}
ggpcoa_hell

```



