---
title: "Analysis_Arkaute"
author: "Javier Porras"
date: "`r Sys.Date()`"
output:
  html_document: default
  pdf_document: default
---
## Dudas 
- Podríamos tener distintos scripts, pero luego "llamar" a estos scripts en distintos chunks? 
- Por qué me aparece ahora en el Knit todas las respuestas de la consola de R? si pone "(echo = FALSE)"

## Cosas qué hacer
- Ir organizando el workflow de este .Rmd. 
- Mirar si hay correlaciones entre alturas y circunferencias para completar los datos de biomasa de muestreos 0, 1 y 2. 
- Sensitivity test de Z
- Rank abundance distribution analysis
- Mirar si podemos utilizar varias ecuaciones en función de la especie. 
- Mirar diferencias en la distribución de los datos de biomasa una vez aplicada la ecuación y después de calcular las biomasas por especie y plot. Lo esperable es que una vez estimada la biomasa en base a la abundancia el histograma sea más "alargado" en la cola.
- Cambiar la forma de visualizar el plot


```{r setup, echo = FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
```

```{r packages}
library(lubridate)
library(stringr)
library(ggplot2)
library(dplyr)
library(reshape2)
library(readr)
library(ggpubr)
library(tidyverse)
library(gridExtra)
source("code/tools/basicFun.R")
```


## Opening and checking flora data

In this script we take a look to the shape of our collected data. 

```{r opening and checking data, include = TRUE}
source("code/codes_rmd_workflow/opening_floradata.R")

```

We have asymmetric data. 

## Transforming data

```{r}
source("code/codes_rmd_workflow/preparing_floradata.R")

```


## Graphs for dynamics

Graph of biomass, diversity and abundance along samplings. Each boxplot contains 4 points (4 replicates per treatment): 

- Biomass: the total biomass (addition of all individual data of biomass at species level)
- Diversity: the total number of species per replicate. 
- Abundance: The addition of all abundance data in order to know the total coverage of each replicate. No species-level information here. 

```{r dynamic graphs}
print(ggarrange(

ggsamplings_biomass <- 
  
ggplot(flora_samplings, aes(x = sampling, y = biomass, color = treatment)) +
  geom_boxplot() +
  labs(x = " ", y = "Biomass") +
  scale_color_manual(values = c("c" = "blue", "p" = "green", "w" = "red", "wp" = "purple")) +
  theme_minimal() +
  geom_vline(xintercept = 1.5, linetype = "dotted", color = "maroon", size = 0.8) +
  theme(legend.position = "none", 
        axis.text = element_text(size = 12)) +
    coord_cartesian(ylim = c(0, 500)),

ggsamplings_diversity <-
  
  ggplot(flora_samplings, aes(x = sampling, y = n_species, color = treatment)) +
  geom_boxplot() +
  labs(x = " ", y = "Diversity") +
  scale_color_manual(values = c("c" = "blue", "p" = "green", "w" = "red", "wp" = "purple")) + 
  geom_vline(xintercept = 1.5, linetype = "dotted", color = "maroon", size = 0.8) +
  theme_minimal() +
  theme(legend.position = "none", 
        axis.text = element_text(size = 12)),

ggsamplings_abundance <- 
 
 ggplot(flora_samplings, aes(x = sampling, y = abundance, color = treatment)) +
  geom_boxplot() +
  labs(x = " ", y = "Abundance") +
  scale_color_manual(values = c("c" = "blue", "p" = "green", "w" = "red", "wp" = "purple")) +
  theme_minimal() +
  theme(legend.position = "none", 
        axis.text = element_text(size = 12)) +
  geom_vline(xintercept = 1.5, linetype = "dotted", color = "maroon", size = 0.8),

labels = c("A", "B", "C"),
ncol = 1, nrow = 3))


```

In the next graph we can see the aggregated differences between treatments: 

```{r Graph for treatments}

print(ggarrange(
  
  ggtreat_biomass <-
  ggplot(flora_samplings, aes(x = treatment, y = biomass, color = treatment)) +
  geom_boxplot() +
  labs(x = " ", y = "Biomass") +
  scale_color_manual(values = c("c" = "blue", "p" = "green", "w" = "red", "wp" = "purple")) +
  theme_minimal() + 
  coord_cartesian(ylim = c(0, 700))+
  theme(legend.position = "none"),

ggtreat_diversity <- 
  ggplot(flora_samplings, aes(x = treatment, y = n_species, color = treatment)) +
  geom_boxplot() +
  labs(x = " ", y = "Diversity") +
  scale_color_manual(values = c("c" = "blue", "p" = "green", "w" = "red", "wp" = "purple")) +
  theme_minimal()+
  theme(legend.position = "none"),

ggtreat_abundance <- 
  ggplot(flora_samplings, aes(x = treatment, y = abundance, color = treatment)) +
  geom_boxplot() +
  labs(x = " ", y = "Abundance") +
  scale_color_manual(values = c("c" = "blue", "p" = "green", "w" = "red", "wp" = "purple")) +
  theme_minimal()+
  theme(legend.position = "none"),

labels = c("A", "B", "C"),
ncol = 3, nrow = 1))

```

